---
title: "Class Project"
author: "Shane Hesprich"
date: "2023-04-28"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Executive Summary



# Introduction

Whether for environmental or economic reasons, fuel efficiency has always been 
an important factor when evaluating a automobile. This paper aims to determines 
if the transmission type (automatic or manual) plays a significant role in an
automobiles fuel efficiency, measured in miles per gallon (mpg).

# Methods/Results

## The Data

To evaluate the role transmission type plays in fuel efficiency we will be using 
the included `R` data set `mtcars`. From `?mtcars` the data set is described as:

> A data frame with 32 observations on 11 (numeric) variables.
> 
> [, 1]	mpg	Miles/(US) gallon
>
> [, 2]	cyl	Number of cylinders
>
> [, 3]	disp	Displacement (cu.in.)
>
> [, 4]	hp	Gross horsepower
>
> [, 5]	drat	Rear axle ratio
>
> [, 6]	wt	Weight (1000 lbs)
>
> [, 7]	qsec	1/4 mile time
>
> [, 8]	vs	Engine (0 = V-shaped, 1 = straight)
>
> [, 9]	am	Transmission (0 = automatic, 1 = manual)
>
> [,10]	gear	Number of forward gears
>
> [,11]	carb	Number of carburetors

```{r}
library(dplyr)
data(mtcars)
# Convert cyl, am, gear, and carb to factor vars
mtcars <- mutate(mtcars, fcyl = factor(cyl), fam = factor(am), fgear = factor(gear), fcarb = factor(carb))
```

## Model Selection

It seems reasonable to assume that horsepower is closely related to 1/4 mile 
time, number of carburetors, and number of cylinders. In order to not over fit 
the model, we will use and ANOVA test to determine which of these variables to 
include in the regression.

```{r}
mdl0 <- lm(mpg ~ fam, data = mtcars)
mdl1 <- lm(mpg ~ fam + hp, data = mtcars)
mdl2 <- lm(mpg ~ fam + hp + fcarb, data = mtcars)
mdl3 <- lm(mpg ~ fam + hp + fcarb + fcyl, data = mtcars)
mdl4 <- lm(mpg ~ fam + hp + fcarb + fcyl + qsec, data = mtcars)

anova(mdl0, mdl1, mdl2, mdl3, mdl4)
```
Additionally, it seems reasonable to conclude that 


## Regression



# Discussion



# Conclusion
