---
title: "Class Project"
author: "Shane Hesprich"
date: "2023-04-28"
output: pdf_document
bibliography: citations.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Executive Summary

This document investigates whether cars with automatic or manual transmissions 
have a lower mpg. A quick exploratory analysis showed automatic having a lower 
mpg than manual (17.15 mpg vs. 24.40mpg, p < 0.01). However, after removing 
unnecessary variables to avoid overfit and a single outlier, a multivariate 
linear regression was performed. This model showed a 1.84 mpg decrease in mpg 
when going from manual to automatic. However, this value was not significant 
(p = 0.35). Based on these results we cannot conclude that transmission type 
has a significant impact on mpg.

# Introduction

Whether for environmental or economic reasons, fuel efficiency has always been 
an important factor when evaluating a automobile. This paper aims to determines 
if the transmission type (automatic or manual) plays a significant role in an
automobiles fuel efficiency, measured in miles per gallon (mpg).

# Methods/Results

## The Data

To evaluate the role transmission type plays in fuel efficiency we will be using 
the included `R` data set `mtcars`. From `?mtcars` the data set is described as:

> A data frame with 32 observations on 11 (numeric) variables.
> 
> [, 1]	mpg	Miles/(US) gallon
>
> [, 2]	cyl	Number of cylinders
>
> [, 3]	disp	Displacement (cu.in.)
>
> [, 4]	hp	Gross horsepower
>
> [, 5]	drat	Rear axle ratio
>
> [, 6]	wt	Weight (1000 lbs)
>
> [, 7]	qsec	1/4 mile time
>
> [, 8]	vs	Engine (0 = V-shaped, 1 = straight)
>
> [, 9]	am	Transmission (0 = automatic, 1 = manual)
>
> [,10]	gear	Number of forward gears
>
> [,11]	carb	Number of carburetors

## Exploratory Analysis

First let's load the data and do some simple plots and statistics to get a feel 
for the data. 
```{r, fig.height=3.5,fig.width=6, message=FALSE}
data(mtcars)
library(dplyr)
library(ggplot2)
# Convert cyl, am, gear, and carb to factor vars
mtcars <- mutate(mtcars, fcyl = factor(cyl), fam = factor(am), 
                 fgear = factor(gear), fcarb = factor(carb), fvs = factor(vs))

g <- ggplot(mtcars, aes(x = fam, y = mpg))
g <- g + geom_boxplot()
g
```
As we can see here, it appears as if automatic transmission (```fam = 0```) has a 
considerably lower median mpg. Let's see if a t-test corroborates what we see 
in the above plot.

```{r}
automatic <- mtcars %>% 
  filter(am == 0) %>%
  select(mpg)
manual <- mtcars %>% 
  filter(am == 1) %>%
  select(mpg)

t.test(manual$mpg, automatic$mpg, alternative = "greater", 
       conf.level = 0.95, paired=FALSE)
```
According to the above t test, manual transmissions have a statistically 
significantly higher mpg than automatic transmissions. Now let's look at a 
linear regression that controls for the other 9 variables and see if this 
observation holds true.

## Model Selection

It seems reasonable to assume that horsepower is closely related to 1/4 mile 
time, number of carburetors, and number of cylinders. In order to not over fit 
the model, we will use and ANOVA test to determine which of these variables to 
include in the regression.

```{r}
mdl0 <- lm(mpg ~ fam, data = mtcars)
mdl1 <- lm(mpg ~ fam + hp, data = mtcars)
mdl2 <- lm(mpg ~ fam + hp + fcarb, data = mtcars)
mdl3 <- lm(mpg ~ fam + hp + fcarb + fcyl, data = mtcars)
mdl4 <- lm(mpg ~ fam + hp + fcarb + fcyl + qsec, data = mtcars)

anova(mdl0, mdl1, mdl2, mdl3, mdl4)
```

## Regression

Let's perform a regression on the remaining variables.

```{r}
fit <- lm(mpg ~ fam + disp + hp + drat + wt + fvs + gear, data = mtcars)
```

## Model evaluation

Before discussing the results of the model, let us first evaluate the model to 
ensure it is a good fit. First we will check the residuals.

```{r}
par(mfrow = c(2, 2))
plot(fit)
```
Next we will look for any outliers. This will be done by looking at the dfbetas 
and hatvalues of our data. According to @belsley2005regression we are looking 
for a $|dfbeta| > \frac{2}{\sqrt{n}}$ and $hatvalue > 2*\left(\frac{n}{p}\right)$, where $n$ is 
the number of samples, and $p$ is the number of estimated parameters. In our 
case $n = 32$ and $p = 8$.

```{r}
n <- dim(mtcars)[1]
p <- 8
dfthresh <- 2/sqrt(n)
hvthresh <- 2*(p/n)
dfbs <- as.data.frame(dfbetas(fit))
dfbs <- dfbs %>% 
  mutate(hv = hatvalues(fit)) %>% 
  filter(if_any('(Intercept)':gear, ~ abs(.) > dfthresh)) %>% 
  filter(hv > hvthresh)
dfbs
```
Let's remove the outlier and rerun the regression.

```{r}
rns <- rownames(dfbs)
mtcars <- mtcars[!(row.names(mtcars) %in% rns),]
fit <- lm(mpg ~ fam + disp + hp + drat + wt + fvs + gear, data = mtcars)
```

# Final Results

Now that we have selected a model, removed outliers, and performed the 
regression let's take a look at the model.

```{r}
summary(fit)$coefficients
```
According to this summary, our model has an adjusted $R^2$ of $0.81$. Based on 
this, our model appears to do a good job predicting mpg. Additionally, our 
model predicts that switch from automatic transmission to manual transmission 
will result in a 1.84 mpg increase. However, with a standard error of 1.934 and 
a t value of 0.953, this value is not statistically significant. Therefore we 
cannot say that there is a statistical difference in the mpgs of manual vs 
automatic transmissions.

# References